---
title: "p8105_hw5_yj2687"
author: "Yijia Jiang"
date: "2022-11-06"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



# Problem 1 (Longitudinal Study)
```{r, warning = FALSE, message=FALSE}

```


&nbsp;

# Problem 2 (Homicide Dataset)
```{r, warning = FALSE, message=FALSE}
# Import dataset
homicide_raw <- read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

* The raw data collected by Washington Post reports a total of `r homicide_raw %>% nrow()` criminal homicides over the past decade in 50 of the largest American cities, including `r homicide_raw %>% ncol()` variables, namely `r homicide_raw %>% names`.
* The reported date, longitude and latitude are designated as numeric variables, and the remaining 9 variables including age, are defined as characters. 
* We can observe some entries for the victim's race, age, and sex are reported as unknown.
* There exist `r sum(is.na(homicide_raw$lat))` missing values in `lat`, `r sum(is.na(homicide_raw$lon))` in `lon`, accounting for `r sum(is.na(homicide_raw$lat)/nrow(homicide_raw)) %>% scales::percent(0.01)`, `r sum(is.na(homicide_raw$lon)/nrow(homicide_raw)) %>% scales::percent(0.01)`, respectively. 
* In particular, there was one record from the city of `Tulsa`, which is located in the state of `OK`, was incorrectly logged as being in the state of `AL`, which will be removed as a typo.


```{r, warning = FALSE, message=FALSE}
# Clean dataset and create variables city_state, resolved
homicide_df = homicide_raw %>% 
  janitor::clean_names() %>%
  mutate(reported_date = as.Date(as.character(reported_date), format = "%Y%m%d")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolved = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest" ~ "unsolved",
      disposition == "Closed by arrest" ~ "solved",
    )) %>% 
  relocate(city_state) %>%
  filter(city_state != "Tulsa, AL")
```


```{r, warning = FALSE, message=FALSE}
# Summarize within cities to calculate the total number of homicides and unsolved homicides
summary_hom_df = homicide_df %>% 
  group_by(city_state) %>% 
  summarize(
    hom_total = n(),
    hom_unsolved = sum(resolved == "unsolved")
  )

summary_hom_df %>% 
  knitr::kable(align = "lrr", 
               col.names = c("City", "Total number of homicides", "Number of unsovled homicides"))
```


```{r, warning = FALSE, message=FALSE}
# Estimate the proportion of unsolved homicides in Baltimore, MD
prop_baltimore = prop.test(
  summary_hom_df %>% filter(city_state == "Baltimore, MD") %>% pull(hom_unsolved), 
  summary_hom_df %>% filter(city_state == "Baltimore, MD") %>% pull(hom_total)) 

prop_baltimore %>% broom::tidy()
```


```{r, warning = FALSE, message=FALSE}
# Iterate to estimate the proportion of unsolved homicides in all cities
prop_cities = 
  summary_hom_df %>% 
  mutate(
    prop_tests = purrr::map2(.x = hom_unsolved, .y = hom_total, ~prop.test(x = .x, n = .y)),
    tidy_tests = purrr::map(.x = prop_tests, ~broom::tidy(.x))
  ) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high)
```



```{r, warning = FALSE, message=FALSE}
# Create a plot showing the estimates and CIs for each city
prop_cities %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Estimates and 95% CIs of Proportion of Unsolved Homicides in All Cities",
    x = "City",
    y = "Estimate")
```



&nbsp;



# Problem 3 (Simulation to Explore Power)
```{r, warning = FALSE, message=FALSE}
# Construct function for one-sample t-test
set.seed(123)
ttest = function(mu){
  sim_data = tibble(x = rnorm(n = 30, mean = mu, sd = 5))
  sim_data %>% 
    summarize(t.test(x, mu = 0, conf.level = 0.95) %>% broom::tidy())
} 


# Generate 5000 datasets from the model and repeat the t-test
sim_result = function(mu){
  rerun(5000, ttest(mu)) %>%
    bind_rows() %>%
    select(p.value, estimate)
}


# mu = 0
sim_result(0)


# mu = 1,2,3,4,5,6


```


```{r, warning = FALSE, message=FALSE}

```

```{r, warning = FALSE, message=FALSE}

```

